---
title: "6252320_Assignment_3"
author: "Leon Debnath"
date: "02/10/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment 3

## Exercise 1

The Beta distribution $Beta(a, b)$ has the following Probability Density Function (PDF):
$$
f(x) =cx^{a-1}(1-x)^{b-1}, \text{ for } 0< x <1,
$$
Using the previous values from the last exercise where: $a = 5,$ $b = 4,$ and $c = 30$ we can form a distribution through rejection sampling:

```{r}
a <- 5
b <- 2
beta <- function(x) 30*x^(5-1) * (1-x)^(2-1)

set.seed(12345) 
n = 10^5
X_rand <- runif(n, min=0, max=5)
Y_rand <- runif(n, min=0, max=5)

beta_dist <- lapply(X_rand, beta) >= Y_rand

X_beta <- X_rand[beta_dist]
Y_beta <- Y_rand[beta_dist]

plot(X_beta, Y_beta, main='Beta Distribution')
```

We sort the list of $Y$ values and pick the central value.  In the case of an even number of values we must take the average of the two central values.

```{r}
sorted <- sort(Y_beta)

# if even -> take the average of the two middle values of the sorted list
if( length(sorted) %% 2 == 0){
  first <- length(sorted)/2
  second <- length(sorted)/2 + 1
  median_est <- (sorted[first] + sorted[second])/2
  
# if odd -> take the middle value of the sorted list
} else {
  middle <- length(sorted)/2
  median_est <- sorted[middle]
}
```

To check the answer, we can use built in functions dbeta and median to calculate the true value.

```{r}
X <- seq(0, 1, by = 0.02)
Y <- dbeta(X, shape1=a, shape2=b)
median_act <- median(Y, na.rm = FALSE)

plot(X,Y)
abline(v=median_est, col='red')
text(0.85, 1, labels='Est', col='red')
abline(v=median_act, col='blue')
text(0.65, 1, labels='Act', col='blue')

sprintf("The estimated median is: %f", median_est)
sprintf("The calculated median is: %f", median_act)
```
## Exercise 2

The Central Limit Theorem (CLT) states that given a sufficiently large sample size, the sampling distribution of the sample means approaches a normal distribution — no matter what the shape of the population distribution.

In the case of a Galton Board the probability of a ball landing in any given slot can be calculated by counting the number of paths that lead there and then dividing that value by the total number of paths. The resulting distribution is discrete and can be calculated by evaluating the Binomial function:
$$
f(k, n) = {}_{n} C_{k} \cdot P(k,n) 
$$
By taking the mean of the binomial we can simulate the final slot that the ball lands in.  Summing the number of times the mean value equals the value of each slot (rounding to the nearest integer) is equivalent to allowing each slot to fill with multiple balls.  Given a large enough sample size, by the CLT the values will form a normal distribution as plotted below

```{r}
balls_dropped <- 100
slots <- 50
balls_per_slot <- rep(0, slots)
for(ball in 1:balls_dropped)
{
  mean <- rbinom(100, slots, 0.5)                    # Calculate the final landing place
  mean <- trunc(mean)                                # Truncate to integer value
  balls_per_slot[mean] <- balls_per_slot[mean] + 1   # Increment the sum for the slot
}
plot(balls_per_slot)
```
