---
title: "Assignment_1"
author: "6252320 Leon Debnath"
date: "10/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 1 (Named Distributions)
Calculate the sum of the probabilities for the total number of tickets sold (80) that each individual attends (85% per person) up to the full capacity of the venue (70 seats) 

```{r}
k <- 70
n <- 80
p <- 0.85

pbinom(q=k, size=n, prob=p)
```
### Question 2 (Independence)
Real world examples of

1. Two random variables that are independent and identically distributed
$$
\text{Let } D \text{ be the value of a rolled die where }D \in \{1,2,3,4,5,6\}\text{, such that the probability }P(D=1)=1/6\\
\text{As each roll of the die is an independent and identically distributed event where } D(s) \text{ is the possible outcome } s \text{ for each roll}\\
D(s_{1}) = D(s_{2})
$$
2. Two random variables that are independent and not identically distributed
$$
\text{Let } D \text{ be the combined value of two rolled dice } D \in \{2 ... 12\} \text{. Despite each die having a probability of landing on any value as before }\\
\text{where }P(D) = 1/6\text{ the combined value has an uneven distribution give the fact that }d_{1} + d_{2} = 1\text{ can never occur, but the value 10 can be}\\
\text{made up of }4 + 6\text{ and }5 + 5.\text{ The intuition behind this can be shown in a histogram of results proving that a random sample of independent variables}\\
\text{are not distributed identically.}
$$
```{r}
D <- c(1,2,3,4,5,6)
D <- rep(D,6)
E <- c(rep(1,6), rep(2,6), rep(3,6), rep(4,6), rep(5,6), rep(6,6)) 
D <- D + E
hist(D, breaks=c(36))
```

3. Two random variables that are dependent and identically distributed


4. Two random variables that are dependent and not identically distributed

